Replication on multicore considering hierarchy and characteristics of the network

### 2013-03-28 paper.tex
Added pseudo-code for how to initialize channels for a fully-meshed network

\documentclass{article}
\usepackage{url,color,xspace,verbatim,subfig,ctable,multirow,listings}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{rotating}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{graphics}
\usepackage{enumitem}
\usepackage{times}
\usepackage{amssymb}
\usepackage[colorlinks=true]{hyperref}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model}

Characteristics of the network influence the performance of
distributed algorithms. Multicores networks are heterogeneous
combinations of complex networks, which makes algorithm selection
hard. %
Furthermore, their diversity makes manual selection tedious. Hence, we
aim for an automatic approach for selecting algorithms for a specific
network. %
We express the network characteristics in a multicore network
model. This model at this stage is by no means complete. We are going to
extend it over time to address more complexities of current and future
multicores.

The network model is defined by communication channels, which in turn
are defined by a set of attributes (here sorted by priority):
\begin{itemize}
\item latency (quantitative) between each pair of nodes (per-edge)
\item ``breakdown'' (quantitative) of message send cost: Ethernet has
  dominating propagation time vs.\ dominating send and receive time on
  multicores (per-node)
\item synchronized clocks: sychronized clocks + bound on propagation
  time $\rightarrow$ \emph{synchronous network} model. \\Then:
  $\exists$ algorithm for synchronous networks $\rightarrow
  \exists$ algorithm for asynchronous
\item reliability (loss, link failure, in-order delivery, phantoms,
  corruption, duplicated messages): Ethernet is not reliable,
  multicore interconnect networks often are
\item Security \& and failure detection (and the cost for that)
\end{itemize}

% Express as:
We express the model in a graph, where nodes are processors and edges
are (on-hop?) communication links. The machines we consider for now
only have bi-directional communication links. Other machines (like the
SCC) have uni-directional links, which is why we model the graph with
directed graph.

% How to deduct tree from model?
We then use this graph to build a group-communication
tree. Interesting techniques could be: clustering, to find closely
connected subsets, distributed messages across cluster first, and then
within cluster. Maybe also a spanning-tree protocol, to find a minimal
graph such that all nodes can be reached.
% Tree -> scheduling
Assuming the tree, there is a scheduling problems for the order in
which to send messages to children. The intuition is to send on
long-delay links first to hide the more expensive latencies.

@@

% EXAMPLE
An example of a machine model, which approximates gruyere can be found
in ..

\begin{tikzpicture}[>=latex,line join=bevel,scale=3.0]
  \pgfsetlinewidth{1bp}
\input{graphs/full_gruyere}
\end{tikzpicture}

>>>>

\documentclass{article}
\usepackage{url,color,xspace,verbatim,subfig,ctable,multirow,listings}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{rotating}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{graphics}
\usepackage{enumitem}
\usepackage{times}
\usepackage{amssymb}
\usepackage[colorlinks=true]{hyperref}
\usepackage[ruled,vlined]{algorithm2e}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The multicore network model}

% selecting hard
Characteristics of the network influence the performance of
distributed algorithms. Due to the heterogeneous combinations of
complex networks on multicores, selection of algorithms is
particularly hard.

% model
We express the network characteristics in a multicore network
model. This model at this stage is by no means complete. We are going to
extend it over time to address more complexities of current and future
multicores.

% graph
We represent the multicore network as a graph. The nodes in this graph
correspond to cores, and edges the communication channels between
nodes. The machines we consider for now only have bi-directional
communication links. Other machines (like the SCC) have
uni-directional links, which is why we model the graph as directed
graph.

% shared memory -> fully meshed
This graph is fully meshed for shared memory machines. Every core can
communicate with every other node via the cache coherence protocol.

\subsection{Node characteristics}

Every node (i.e.\ core) in the system has characteristics associated
with it:

\begin{description}
\item[synchrony] Parts of a multicore machine are synchronous. Hence,
  one characteristic of a node is an identifier for the synchrony
  cluster of a node.
\item[$t_{send}$] Time required to send a message, i.e.\ marshalling
  and queuing the message
\item[$t_{recv}$] Time required to receive a message, i.e.\
  dequeuing, unmarshalling and demultiplexing a channel. 
\end{description}

\subsection{Link characteristics}

The network model is defined by communication channels, which in turn
are defined by a set of attributes (here sorted by priority):

\begin{description}
\item[latency] (quantitative) between each pair of nodes (per-edge)
\item[breakdown] (quantitative) of message send cost: Ethernet has
  dominating propagation time vs.\ dominating send and receive time on
  multicores (per-node)
\item[bounded delivery]
\item[reliability] (loss, link failure, in-order delivery, phantoms,
  corruption, duplicated messages): Ethernet is not reliable,
  multicore interconnect networks often are
\item[Security] \& and failure detection (and the cost for that)
\end{description}

@@

\subsection{Acquiring the model}

\subsection{Example}

An example of a machine model, which approximates a 2x2x1 nos box can
be found in Figure~\ref{fig:full_nos}

\begin{figure}
\begin{tikzpicture}[>=latex,line join=bevel]
  \pgfsetlinewidth{1bp}
\input{graphs/full_nos}
\end{tikzpicture}
\caption{Network model of a 2x2x1 machine}
\label{fig:full_nos}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Example: Atomic broadcast
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Atomic Broadcast}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Building the tree}

First, we build the broadcast tree. We kick-start our protocol
connecting every process with all other processes to get a
fully-meshed network of channels.

We use a round-based algorithm to open channels. Then, every node
knows exactly the source of an incoming connection. We formalize this
algorithm in Algorithm~\ref{algo:ab_bind}.

\begin{algorithm}[H]
  %
  \SetKwInOut{Assumptions}{assumptions}
  \Assumptions{Processes have unique contiguous
    identifiers starting at 0}
  \BlankLine
  %
  \SetKwArray{c}{channels}
  \SetKwFunction{connectNode}{connectNode}
  \SetKwFunction{listen}{listen}
  % 
  \KwData{process id $p$, round $r$, %
    each process an array of channels \c}
  \KwResult{Fully-meshed network of processes}
  % 
  \BlankLine
  %
  $r \leftarrow 0$\;
  \For{$i \leftarrow 0$ \KwTo $num(p)$}{
    \eIf{$p=i$}{
      \For{$o \leftarrow i+1$ \KwTo $num(p)$}{
        \c{$o$} $\leftarrow$ \connectNode{$o$}\;
      }
    }{
        \c{$i$} $\leftarrow$ \listen{}\;
    }
  }
  \caption{Atomic broadcast on reliable communication channels}
  \label{algo:ab}
\end{algorithm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Broadcast protocol}

% Our implementation
We now build an atomic broadcast for multicore machines. Our
implementation is based on the Barrelfish UMP interconnect driver. It
provides reliable channels with in-order delivery. Communication
starts at the root node, which acts as a
sequentializer. See~\ref{algo:ab} for details.

\begin{algorithm}[H]
  \SetAlgoLined
  %
  \KwData{List of processes $p$, broadcast tree as graph $(V, E)$}
  \KwResult{Tree based atomic broadcast using a sequentializer}
  % 
  $(V, E)$ = build\_graph\;
  \While{not at end of this document}{
    read current\;
    \eIf{understand}{
      go to next section\;
      current section becomes this one\;
    }{
      go back to the beginning of current section\;
    }
  }
  \caption{Atomic broadcast on reliable communication channels}
  \label{algo:ab}
\end{algorithm}

% How to deduct tree from model?
We then use this graph to build a group-communication
tree. Interesting techniques could be: clustering, to find closely
connected subsets, distributed messages across cluster first, and then
within cluster. Maybe also a spanning-tree protocol, to find a minimal
graph such that all nodes can be reached.

% Tree -> scheduling
Assuming the tree, there is a scheduling problems for the order in
which to send messages to children. The intuition is to send on
long-delay links first to hide the more expensive latencies.
